---
title: "Homework_21.qmd"
format: html
editor: visual
---

```{r}
# Setting up R 
rm(list=ls())
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
```

Chapter 5: Introducing stats in R

X\^2 contingency table analysis is an analysis of count data. Test of association of two or more categorical variables.

```{r}
# read in data
lady <- read.csv(here("Data", "ladybirds.csv"))

# check it out
glimpse(lady)
```

Organizing data for plotting and analysis

```{r}
# organizing
totals <- lady |> 
  group_by(Habitat, colour) |> 
  summarise(total.number = sum(number))

# bar chart to summarise data
ggplot(totals, aes(x = Habitat, y = total.number, 
                   fill = colour)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_manual(values = c(black = "black", red = "red"))
```

identity() = tells ggplot not to calculate anything from data

position = "dodge" pulls bars next to one another

xtabs = transforms data frame into a matrix

```{r}
# cross tabulate # of columns of counts in totals data frame by these variables
lady.mat <- xtabs(number ~ Habitat + colour, 
                  data = lady)
lady.mat
```

```{r}
lady.chi <- chisq.test(lady.mat)
names(lady.chi)

lady.chi$expected
```

Two sample T-Test: Comparision of means of two groups of numeric values. Appripriate when sample sizes of both groups are small. Assumes data in each group are normally distributed and their variances are equal.

```{r}
# read in data
ozone <- read.csv(here("Data", "ozone.csv"))

# check it out
glimpse(ozone)
```

```{r}
# Plot the data
ggplot(ozone, aes(x = Ozone)) + 
  geom_histogram(binwidth = 10) + 
  facet_wrap(~ Garden.location, ncol = 1) + 
  theme_bw()
```

```{r}
# t test
t.test(Ozone ~ Garden.location, data = ozone)

var.test(Ozone ~ Garden.location, data = ozone)
```

Do ozone levels (Ozone) vary as a function of location (Garden.location)
