---
title: "Homework_23"
format: html
editor: visual
---

```{r}
# clear R's brain
rm(list=ls())
library(tidyverse)
library(dplyr)
library(ggplot2)
library(here)
```

```{r}
# read in csv file
plant_gr <- read.csv(here("Data", "plant.growth.rate.csv"))

glimpse(plant_gr)
```

```{r}
# make figure of data
ggplot(plant_gr, 
       aes(x = soil.moisture.content, y = plant.growth.rate)) +
  geom_point() +
  xlab("Plant Growth Rate (mm/week)") +
    theme_bw()
```

```{r}
# making simple linear regression
model_pgr <- lm(plant.growth.rate ~ soil.moisture.content, 
                data = plant_gr)

# reads as 'fit a linear model where we hypothosize that plant growth rate is a function of soil moisture content, using the variables from the plant_gr dataframe
```

```{r}
# check assumptions and plot
library(ggfortifiy)
autoplot(model_pgr, smooth.colour = NA)
```

1.  Top left graph: "systematic part of the model. it tells us whether a line is appropriate to fit in the data. If things are wrong, hump shapes or valleys will be apparent, meaning that the structure of the model is wrong.
2.  Top Right: This evaluates the assumption of normality of the residuals. The dots are the residuals and the dashed line is the expectation under a normal distribution.
3.  Bottom left: This evaluates the assumption of equal variance. The y-axis is standardized indicator of the variation. Linear models assume that the variance is constant over all predicated values of the response variable. There should be no pattern.
4.  Bottom right: This evaluates leverage, a tool not only to detect influential data points, ones that move the gradient more than might be expected, but also to detect outliers.

```{r}
# interpretation
anova(model_pgr)
summary(model_pgr)
```

```{r}
ggplot(plant_gr, aes(x = soil.moisture.content, 
                     y = plant.grwoth.rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ylab("Plant Growth Rate (mm/week)") +
  theme_bw()
```
