---
title: "Homework 24"
Name: Alexis Parent 
Date: 10/31/2024
format: html
editor: visual
---

## Overview

For your homework you will implement a data analysis using a dataset on crab growth measurements.

Set things up (hint: there is a new package to add with `library()`)

```{r}
# set up R and clear mind 
rm(list=ls())
library(tidyverse)
library(ggplot2)
library(ggfortify)
library(here)
```

## The data

Let's pull in some data appropriate biologists: crab measurements. This dataset includes measurements of farmed crabs.

```{r}
# read in data and take a look
crabs <- read.csv(here("Data", "CrabAgePrediction.csv"))
glimpse(crabs)
view(crabs)
```

One of the variables is categorical - let's fix that in the data.

```{r}
# chnage variables to factor
crabs$Sex<- as.factor(crabs$Sex)
levels(crabs$Sex)
```

## Question

Is there a relationship between height and weight in female crabs?

1.  Extract the data for just female crabs.

    ```{r}
    # take data from just female crabs 
    female_crabs <- crabs %>% filter(Sex == "F") 

    view(female_crabs)
    ```

2.  Examine the response and predictor variables and use the flow chart from class to decide the appropriate statistical analysis (choose from: chi-square, 2-sample t-test or linear regression). Explain your answer here:

    ```{r}
    # The response variable is height and weight. 
    # The predicator variable is if the crab is male of female. 
    # The appropriate statistical analysis would be a chi-square test,    because both crab height and weight are ratio data. Furthermore, there are also no more than two groups to compare. 
    ```

3.  Work through the data analysis **following all of the steps provided in chapter 5 of your book**. You may use class notes or homework keys as a guide.

    ```{r}
    female_crabs %>% 
      group_by(Height, Weight)

    glimpse(female_crabs)

    # make a figure 
    ggplot(female_crabs, aes(x = factor(Height), y = Weight)) +
     geom_bar(stat = "identity", fill = "blue") 

    # X2 contingancy table analysis
    female_crabs.mat <- xtabs(Weight ~ Height, 
                              data = female_crabs)
      
    ```

```{r}
# Conduct test!!
chisq.test(female_crabs.mat)

# Return values
female_crabs.chi <- chisq.test(female_crabs.mat)
names(female_crabs.chi)
```
